---
title: "JOB"
output: html_notebook
---


```{r}
#FIRST RUN THIS CHUNK FOR THE REST TO RUN

library(tidyverse)
library(tidyverse)

load("/Users/jessiechen/Downloads/ICPSR_36692 5/DS0007/36692-0007-Data.rda")
d=da36692.0007  ##{da36692.0007}this data must be in youre local environment

skimr::skim(d)
d %>%  View()

```



```{r}
#MEAT
d %>% select(starts_with("D4MEAT2"))

#EGGS
d %>% select(starts_with("D4EGG2"))

#FOOD MADE FROM BEANS
d %>% select(starts_with("D4BEAN2"))
#SALT PRESERVED
d %>% select(starts_with("D4VEG2"))
#TEA
d %>% select(starts_with("D4TEA2"))

#GARLIC
d %>% select(RD4GARLIC2)
```


#AGE
```{r}
#AGE...A3

age=d  %>%  group_by(TRUEAGE) %>% tally() ##

summary(d$TRUEAGE)
sd(d$TRUEAGE)

###table(d$TRUEAGE)

barplot(table(d$TRUEAGE),main = "AGE BARPLOT",xlab = "AGES")


```

#GENDER
```{r}
#GENDER....A1

gender=d %>% select(A1) %>% group_by(A1) %>% tally()
gender
tablegender <- table(gender)
prop.table(tablegender)


```

#MARITUAL STATUS
```{r}
#MARITAL STATUS....F41

d %>% group_by(F41) %>% select(F41)%>% tally()

marriage=d$F41 %>% fct_collapse(
  married=c("(1) currently married and living with spouse","(2) married but not living with spouse"),
 divorced=c("(3)  divorced","(4) widowed"),
 nevermarried="(5) never married" ) 

table(marriage)
 barplot(table(marriage),main = "MARITUAL STATUS")
summary(marriage)
prop.table(table(marriage))
 
```


#EDUCATION
```{r}
#EDUCATION
# LESS THAn HIGHSCHOOL=year< 12
#highchool=13<years<17
#more than highschool=years>18


lessthan_highschool=d%>%select(F1) %>%  filter(F1<=12) %>% tally();lessthan_highschool

highschool=d%>%select(F1) %>%  filter(13<F1 & F1<=17) %>% tally();highschool

morethan_highschool=d%>%select(F1) %>%  filter(F1>=18) %>% tally();morethan_highschool 

summary(d$F1)
sd(d$F1)


```


#SMOKING
```{r}

smoooking=d %>%group_by(D71) %>%  select(D71) %>% tally()
smoooking

barplot(table(d$D71),ylim = c(0,6000),main = "SMOKKING")
```



#DRINKING
```{r}
#DRINKING
drinking=d %>%group_by(D81) %>%  select(D81) %>% group_by(D81) %>% tally() %>% na.omit()
table(d$D81)
barplot(table(d$D81),main = "DRINKING")
```


#BIVARIATE FRUIT CANCER
```{r}
fruitcancer=d %>% select(G15L1,D31)  %>% 
  mutate(cancer=fct_collapse(G15L1,
  "yes"="(1) yes",
  "no"=c("(2) no","(8) do not know","(9)  missing"   )),
    fruits=  fct_collapse(D31,
              "no"=c("(4) rarely or never", "(8) don't know" ,"(9) missing"  ),
    "yes"=c("(1) almost everyday" ,"(2) quite often","(3) occasionally") ) ) %>%
  select(fruits,cancer) %>% 
  na.omit() 

table(fruitcancer)
prop.table(table(fruitcancer))


chi2 = chisq.test(table(fruitcancer), correct=F)
c(chi2$statistic, chi2$p.value)

```





#BIVARIATE VEG CANCER
```{r}
vegscancer=d %>% select(G15L1,D32)  %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    vegs=  fct_collapse(D32,
                no=c("(4) rarely or never","(8) don't know", "(9) missing"),
     yes=c("(1) almost everyday" ,"(3) occasionally","(2) except winter")) ) %>%
               select(vegs,cancer) %>% na.omit() 

table(vegscancer)
prop.table(table(vegscancer))

```



#BIVARIATE MEAT CANCER
```{r}
meatcancer=d %>% select(G15L1,D4MEAT2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    meat=fct_collapse(D4MEAT2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,meat) %>% na.omit()
  
table(meatcancer)
prop.table(table(meatcancer))

```



#BIVARIATE EGG CANCER

```{r}
eggcancer=d %>% select(G15L1,D4EGG2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    egg=fct_collapse(D4EGG2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(egg,cancer) %>% na.omit()
  

table(eggcancer)
prop.table(table(eggcancer))
```



#BIVARIATE BEAN CANCER

```{r}
beancancer=d %>% select(G15L1,D4BEAN2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    bean=fct_collapse(D4BEAN2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(bean,cancer) %>% na.omit()
  

table(beancancer)
prop.table(table(beancancer))
```



#BIVARIATE SALT PRESERVED CANCER

```{r}
saltcancer=d %>% select(G15L1,D4VEG2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    salt=fct_collapse(D4VEG2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(salt,cancer) %>% na.omit()
  

table(saltcancer)
prop.table(table(saltcancer))
```



#BIVARIATE TEA CANCER

```{r}
teacancer=d %>% select(G15L1,D4TEA2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
  tea=fct_collapse(D4TEA2,no=c("(5) rarely or never" ,"(8) don't know","(6) 6" ),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(tea,cancer) %>% na.omit()
  

table(teacancer)
prop.table(table(teacancer))
```



#BIVARIATE GARLIC CANCER

```{r}
garliccancer=d %>% select(G15L1,D4GARL2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    garlic=fct_collapse(D4GARL2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(garlic,cancer) %>% na.omit()
  

table(garliccancer)
prop.table(table(garliccancer))

```
fish
```{r}
fish=d %>% select(G15L1,D4FISH1) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    fish=fct_collapse(D4FISH1,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(fish,cancer) %>% na.omit()
  

table(fish)
prop.table(table(fish))
```
stable food
```{r}
stablefood=d %>% select(G15L1,D1)%>% 
mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    stablefood=fct_collapse(D1,no=c("(5) rice" ,"(8) corn"),
                      yes=c("(1) wheat","(2) rice and wheat","(9) missing","(3) other"))) %>% 
    select(stablefood,cancer) %>% na.omit()
table(d %>% select(G15L1,D1))

```
```{r}
milk=d %>% select(G15L1,D4MILK2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    milk=fct_collapse(D4MILK2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(milk,cancer) %>% na.omit()
  

table(milk)
prop.table(table(milk))
```
```{r}
nuts=d %>% select(G15L1,D4NUT2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    nuts=fct_collapse(D4NUT2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(nuts,cancer) %>% na.omit()
  

table(nuts)
prop.table(table(nuts))

```
```{r}
mushroom=d %>% select(G15L1,D4ALGA2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    mushroom=fct_collapse(D4ALGA2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(mushroom,cancer) %>% na.omit()
  

table(mushroom)
prop.table(table(mushroom))
```

```{r}
VIT=d %>% select(G15L1,D4VIT2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
   VIT=fct_collapse(D4VIT2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(VIT,cancer) %>% na.omit()
  

table(VIT)
prop.table(table(VIT))
```
```{r}
drug=d %>% select(G15L1,D4DRUG2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
   drug=fct_collapse(D4DRUG2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once per week","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(drug,cancer) %>% na.omit()
  

table(drug)
prop.table(table(drug))
```
```{r}
teatype=d %>% select(G15L1,D4B)%>% 
mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    teatype=fct_collapse(D48,no=c("(5) rice" ,"(8) corn"),
                      yes=c("(1) wheat","(2) rice and wheat","(9) missing","(3) other"))) %>% 
    select(teatype,cancer) %>% na.omit()
table(d %>% select(G15L1,D48))


```
```{r}
sugar=d %>% select(G15L1,D4SUGA1) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
   sugar=fct_collapse(D4SUGA1,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2)  not everyday, but at least once  per week ","(9) missing","(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(sugar,cancer) %>% na.omit()
  

table(sugar)
prop.table(table(sugar))

```


##||


#FRUIT CANCER LOGISTIC
```{r}

fruitcancer=d %>% select(G15L1,D31)  %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    fruits=  fct_collapse(D31,
                no=c("(4) rarely or never", "(8) don't know" ,"(9) missing"  ),
      yes=c("(1) almost everyday" ,"(2) quite often","(3) occasionally") ) ) %>%
               
  select(cancer,fruits) %>% 
  na.omit() 

table(fruitcancer)

summary(glm(fruitcancer$cancer~fruitcancer$fruit,family = "binomial"))

```



#VEGETABLE CANCER LOGISTIC
```{r}
vegscancer=d %>% select(G15L1,D32)  %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    vegs=  fct_collapse(D32,
                no=c("(4) rarely or never","(8) don't know", "(8) don't know"),
     yes=c("(1) almost everyday" ,"(3) occasionally") ,
     rarely="(2) except winter") ) %>%
               
  select(cancer,vegs) %>% 
  na.omit() 

table(vegscancer)

barplot(table(vegscancer))

summary(glm(vegscancer$cancer~vegscancer$vegs,family = "binomial"))

```





#MEAT CANCER LOGISTIC
```{r}

meatcancer=d %>% select(G15L1,D4MEAT2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    meat=fct_collapse(D4MEAT2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,meat) %>% na.omit()
  
table(meatcancer)
barplot(table(meatcancer))
  
summary(glm(meatcancer$cancer~meatcancer$meat,family = "binomial"))

```





#EGG CANCER LOGISTIC
```{r}
eggcancer=d %>% select(G15L1,D4EGG2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    egg=fct_collapse(D4EGG2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,egg) %>% na.omit()
  

table(eggcancer)
barplot(table(eggcancer))

summary(glm(eggcancer$cancer~eggcancer$egg,family = "binomial"))

```







#FOOD PRESERVED BY BEANS AND CANCER LOGISTIC
```{r}
beancancer=d %>% select(G15L1,D4BEAN2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    bean=fct_collapse(D4BEAN2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,bean) %>% na.omit()
  

table(beancancer)
barplot(table(beancancer))

summary(glm(beancancer$cancer~beancancer$bean,family = "binomial"))

```



#SALTY PRESERVED FOOD AND CANCER LOGISTIC

```{r}
saltcancer=d %>% select(G15L1,D4VEG2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    salt=fct_collapse(D4VEG2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,salt) %>% na.omit()
  

table(saltcancer)

summary(glm(saltcancer$cancer~saltcancer$salt,family = "binomial"))

```






#TEA AND CANCER LOGISTIC
```{r}
teacancer=d %>% select(G15L1,D4TEA2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    tea=fct_collapse(D4TEA2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,tea) %>% na.omit()
  

table(teacancer)
barplot(table(teacancer))

summary(glm(teacancer$cancer~teacancer$tea,family = "binomial"))

```






#GARLIC AND CANCER LOGISTIC
```{r}
garliccancer=d %>% select(G15L1,D4GARL2) %>% 
  mutate(cancer=fct_collapse(G15L1,
  yes="(1) yes",
  no=c("(2) no","(8) do not know","(9)  missing"   )),
    garlic=fct_collapse(D4GARL2,no=c("(5) rarely or never" ,"(8) don't know"),
                      yes=c("(1) almost everyday","(2) not everyday, but at least once  per week","(9) missing"),
                      rarely=c("(3) not every week, but at least once  per month","(4) not every month, but occasionally"))) %>% 
    select(cancer,garlic) %>% na.omit()
  

table(garliccancer)
barplot(table(garliccancer))

summary(glm(garliccancer$cancer~garliccancer$garlic,family = "binomial"))

```




